Supplementary material

### 2022-02-09 submission/manuscript.rmd
new updates

In line with previous analyses [@Califf1997; @Steyerberg2000], we fitted a
logistic regression model with 6 baseline covariates, i.e. age, Killip class,
systolic blood pressure, heart rate, an indicator of previous MI, and the
location of MI, to predict 30-day mortality risk. A constant effect of treatment
was included in the model. When deriving risk predictions for individuals we set
the treatment indicator to 0. More information on model development can be found
in the supplement (Section 6).

@@

When assuming a true constant relative treatment effect, discrimination for
benefit was only slightly lower for the linear interaction model, but
substantially lower for the non-linear RCS approaches (Figure
\ref{fig:discrimination}; panel A). With strong linear or quadratic deviations
from a constant relative treatment effect, all methods discriminated quite
similarly (Figure \ref{fig:discrimination}; panels B-C). In the scenario with
non-monotonic deviations, the constant effect model had much lower
discriminative ability compared to all other methods (median AUC of
`r aucDistribution %>% filter(scenarioId == 397) %>% mutate(across(-one_of("scenarioId"), ~ d(.x))) %>% pull(median_constant)` 
for the constant effects model,
`r aucDistribution %>% filter(scenarioId == 397) %>% mutate(across(-one_of("scenarioId"), ~ d(.x))) %>% pull(median_linear)`
for the linear interaction model and
`r aucDistribution %>% filter(scenarioId == 397) %>% mutate(across(-one_of("scenarioId"), ~ d(.x))) %>% pull(median_rcs_3)`
for the best-performing RCS-3; Figure \ref{fig:discrimination}; panel D). The
adaptive approach was unstable in terms of discrimination for benefit,
especially in the presence of treatment-related harms. With increasing number of
RCS knots, we observed decreasing median values and increasing variability of
the c-for-benefit in all scenarios. When we increased the sample size to 17,000
we observed similar trends, however the performance of all methods was more
stable (Supplement, Figure S6). Finally, when we increased the true prediction
AUC to 0.85 the adaptive approach in the case of non-monotonic deviations was,
again, more conservative, especially with null or moderate treatment-related
harms (Supplement, Figure S7).

@@

In terms of calibration for benefit, the constant effects model outperformed all
other models in the scenario with true constant treatment effects, but was
miscalibrated for all deviation scenarios (Figure \ref{fig:calibration}). The
linear interaction model showed best or close to best calibration across all
scenarios and only showed worse calibration compared to RCS-3 in case of
non-monotonic deviations and treatment-related harms (Figure
\ref{fig:calibration}; panel D). The adaptive approach was worse calibrated in
scenarios with strong linear and non-monotonic deviations compared to the linear
interaction model and RCS-3. When we increased sample size to 17,000 similar
conclusions on calibration for benefit could be drawn. As expected, all methods
displayed more stable calibration performance due to the larger number of
patients (Supplement, Figure S8). When we increased the true prediction AUC to
0.85, the linear interaction model was worse calibrated, on average, than RCS-3
in the case of strong quadratic deviations from constant relative treatment
effects (Supplement, Figure S9).

@@

Risk-based approaches to predictive HTE estimate treatment benefit as a function
of baseline risk. A limitation of our study is that we assumed treatment benefit
to be a function of baseline risk in the majority of the simulation
scenarios. We also considered constant moderate and strong treatment-related
harms, applied on the absolute scale to expand the range of scenarios in line
with previous work [@Glasziou1995]. In a limited set of scenarios where we
assumed the existence of true treatment-covariate interactions, our conclusions
remained unchanged. Even though the average error rates increased for all the
considered methods, due to the miss-specification of the outcome model, the
linear interaction model had the lowest error rates. RCS-3 had very comparable
performance. The constant treatment effect model often gave biased results,
espcially in the presence of moderate or strong treatment-related harms. All the
results of these simulations can be found in Supplement, Section 6. Future
simulation studies could explore the effect of more extensive deviations from
risk-based treatment effects.

>>>>

In line with previous analyses [@Califf1997; @Steyerberg2000], we fitted a
logistic regression model with 6 baseline covariates, i.e. age, Killip class,
systolic blood pressure, heart rate, an indicator of previous MI, and the
location of MI, to predict 30-day mortality risk. A constant effect of treatment
was included in the model. When deriving risk predictions for individuals we set
the treatment indicator to 0. More information on model development can be found
in the supplement (Supplement, Section 8).

@@

When assuming a true constant relative treatment effect, discrimination for
benefit was only slightly lower for the linear interaction model, but
substantially lower for the non-linear RCS approaches (Figure
\ref{fig:discrimination}; panel A). With strong linear or quadratic deviations
from a constant relative treatment effect, all methods discriminated quite
similarly (Figure \ref{fig:discrimination}; panels B-C). In the scenario with
non-monotonic deviations, the constant effect model had much lower
discriminative ability compared to all other methods (median AUC of
`r aucDistribution %>% filter(scenarioId == 397) %>% mutate(across(-one_of("scenarioId"), ~ d(.x))) %>% pull(median_constant)` 
for the constant effects model,
`r aucDistribution %>% filter(scenarioId == 397) %>% mutate(across(-one_of("scenarioId"), ~ d(.x))) %>% pull(median_linear)`
for the linear interaction model and
`r aucDistribution %>% filter(scenarioId == 397) %>% mutate(across(-one_of("scenarioId"), ~ d(.x))) %>% pull(median_rcs_3)`
for the best-performing RCS-3; Figure \ref{fig:discrimination}; panel D). The
adaptive approach was unstable in terms of discrimination for benefit,
especially in the presence of treatment-related harms. With increasing number of
RCS knots, we observed decreasing median values and increasing variability of
the c-for-benefit in all scenarios. When we increased the sample size to 17,000
we observed similar trends, however the performance of all methods was more
stable (Supplement, Figure S6). Finally, when we increased the true prediction
AUC to 0.85 the adaptive approach in the case of non-monotonic deviations was,
again, more conservative, especially with null or moderate treatment-related
harms (Supplement, Figure S5).

@@

In terms of calibration for benefit, the constant effects model outperformed all
other models in the scenario with true constant treatment effects, but was
miscalibrated for all deviation scenarios (Figure \ref{fig:calibration}). The
linear interaction model showed best or close to best calibration across all
scenarios and only showed worse calibration compared to RCS-3 in case of
non-monotonic deviations and treatment-related harms (Figure
\ref{fig:calibration}; panel D). The adaptive approach was worse calibrated in
scenarios with strong linear and non-monotonic deviations compared to the linear
interaction model and RCS-3. When we increased sample size to 17,000 similar
conclusions on calibration for benefit could be drawn. As expected, all methods
displayed more stable calibration performance due to the larger number of
patients (Supplement, Figure S6). When we increased the true prediction AUC to
0.85, the linear interaction model was worse calibrated, on average, than RCS-3
in the case of strong quadratic deviations from constant relative treatment
effects (Supplement, Figure S7).

@@

Risk-based approaches to predictive HTE estimate treatment benefit as a function
of baseline risk. A limitation of our study is that we assumed treatment benefit
to be a function of baseline risk in the majority of the simulation
scenarios. We also considered constant moderate and strong treatment-related
harms, applied on the absolute scale to expand the range of scenarios in line
with previous work [@Glasziou1995]. In a limited set of scenarios where we
assumed the existence of true treatment-covariate interactions, our conclusions
remained unchanged. Even though the average error rates increased for all the
considered methods, due to the miss-specification of the outcome model, the
linear interaction model had the lowest error rates. RCS-3 had very comparable
performance. The constant treatment effect model often gave biased results,
espcially in the presence of moderate or strong treatment-related harms. All the
results of these simulations can be found in Supplement, Section 7. Future
simulation studies could explore the effect of more extensive deviations from
risk-based treatment effects.

### 2022-02-09 submission/manuscript.tex
new updates

In line with previous analyses {[}13,14{]}, we fitted a logistic
regression model with 6 baseline covariates, i.e.~age, Killip class,
systolic blood pressure, heart rate, an indicator of previous MI, and
the location of MI, to predict 30-day mortality risk. A constant effect
of treatment was included in the model. When deriving risk predictions
for individuals we set the treatment indicator to 0. More information on
model development can be found in the supplement (Section 6).

@@

When assuming a true constant relative treatment effect, discrimination
for benefit was only slightly lower for the linear interaction model,
but substantially lower for the non-linear RCS approaches (Figure
\ref{fig:discrimination}; panel A). With strong linear or quadratic
deviations from a constant relative treatment effect, all methods
discriminated quite similarly (Figure \ref{fig:discrimination}; panels
B-C). In the scenario with non-monotonic deviations, the constant effect
model had much lower discriminative ability compared to all other
methods (median AUC of 0.4971 for the constant effects model, 0.5285 for
the linear interaction model and 0.5304 for the best-performing RCS-3;
Figure \ref{fig:discrimination}; panel D). The adaptive approach was
unstable in terms of discrimination for benefit, especially in the
presence of treatment-related harms. With increasing number of RCS
knots, we observed decreasing median values and increasing variability
of the c-for-benefit in all scenarios. When we increased the sample size
to 17,000 we observed similar trends, however the performance of all
methods was more stable (Supplement, Figure S6). Finally, when we
increased the true prediction AUC to 0.85 the adaptive approach in the
case of non-monotonic deviations was, again, more conservative,
especially with null or moderate treatment-related harms (Supplement,
Figure S7).

@@

Risk-based approaches to predictive HTE estimate treatment benefit as a
function of baseline risk. A limitation of our study is that we assumed
treatment benefit to be a function of baseline risk in the majority of
the simulation scenarios. We also considered constant moderate and
strong treatment-related harms, applied on the absolute scale to expand
the range of scenarios in line with previous work {[}15{]}. In a limited
set of scenarios where we assumed the existence of true
treatment-covariate interactions, our conclusions remained unchanged.
Even though the average error rates increased for all the considered
methods, due to the miss-specification of the outcome model, the linear
interaction model had the lowest error rates. RCS-3 had very comparable
performance. The constant treatment effect model often gave biased
results, espcially in the presence of moderate or strong
treatment-related harms. All the results of these simulations can be
found in Supplement, Section 6. Future simulation studies could explore
the effect of more extensive deviations from risk-based treatment
effects.

>>>>

In line with previous analyses {[}13,14{]}, we fitted a logistic
regression model with 6 baseline covariates, i.e.~age, Killip class,
systolic blood pressure, heart rate, an indicator of previous MI, and
the location of MI, to predict 30-day mortality risk. A constant effect
of treatment was included in the model. When deriving risk predictions
for individuals we set the treatment indicator to 0. More information on
model development can be found in the supplement (Supplement, Section
8).

@@

When assuming a true constant relative treatment effect, discrimination
for benefit was only slightly lower for the linear interaction model,
but substantially lower for the non-linear RCS approaches (Figure
\ref{fig:discrimination}; panel A). With strong linear or quadratic
deviations from a constant relative treatment effect, all methods
discriminated quite similarly (Figure \ref{fig:discrimination}; panels
B-C). In the scenario with non-monotonic deviations, the constant effect
model had much lower discriminative ability compared to all other
methods (median AUC of 0.4971 for the constant effects model, 0.5285 for
the linear interaction model and 0.5304 for the best-performing RCS-3;
Figure \ref{fig:discrimination}; panel D). The adaptive approach was
unstable in terms of discrimination for benefit, especially in the
presence of treatment-related harms. With increasing number of RCS
knots, we observed decreasing median values and increasing variability
of the c-for-benefit in all scenarios. When we increased the sample size
to 17,000 we observed similar trends, however the performance of all
methods was more stable (Supplement, Figure S6). Finally, when we
increased the true prediction AUC to 0.85 the adaptive approach in the
case of non-monotonic deviations was, again, more conservative,
especially with null or moderate treatment-related harms (Supplement,
Figure S5).

@@

Risk-based approaches to predictive HTE estimate treatment benefit as a
function of baseline risk. A limitation of our study is that we assumed
treatment benefit to be a function of baseline risk in the majority of
the simulation scenarios. We also considered constant moderate and
strong treatment-related harms, applied on the absolute scale to expand
the range of scenarios in line with previous work {[}15{]}. In a limited
set of scenarios where we assumed the existence of true
treatment-covariate interactions, our conclusions remained unchanged.
Even though the average error rates increased for all the considered
methods, due to the miss-specification of the outcome model, the linear
interaction model had the lowest error rates. RCS-3 had very comparable
performance. The constant treatment effect model often gave biased
results, espcially in the presence of moderate or strong
treatment-related harms. All the results of these simulations can be
found in Supplement, Section 7. Future simulation studies could explore
the effect of more extensive deviations from risk-based treatment
effects.

### 2022-02-09 submission/supplement.rmd
new updates

In this section we present specific scenarios from our simulation settings
in which evolution of benefit followed similar patterns to (REF!!). In this case
patients were stratified into risk quarters based on their true baseline
risk. Within each risk quarter we constructed boxplots of true benefit.

@@

```{r, cache=TRUE, echo=FALSE, fig.cap="Simulation scenarios that closely follow trials. In this case, we see increasing absolute benefits with increasing baseline risk.", fig.show="hold", out.width = '50%'}
include_graphics(
  c(
    here("figures/scenario_406.png"),
    here("figures/scenario_422.png")
  )
)
```

@@

>>>>

In this section we present specific scenarios from our simulation settings in
which evolution of benefit followed similar patterns to [@Kent2016]. In this
case patients were stratified into risk quarters based on their true baseline
risk. Within each risk quarter we constructed boxplots of true benefit.

@@

```{r, cache=TRUE, echo=FALSE, fig.cap="Simulation scenarios that closely follow trials. In this case, we see increasing absolute benefits with increasing baseline risk up to the third risk quarter. In the fourth risk quarter this trend is interrupted and benefits are diminished.", fig.show="hold", out.width = '50%'}
include_graphics(
  c(
    here("figures/scenario_406.png"),
    here("figures/scenario_422.png")
  )
)
```

@@

\newpage
# References
\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}
\noindent
<div id="refs"></div>
\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}
\noindent
